---
title: "Elvira Khwatenge programming test"
output: html_document
date: "2024-04-17"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readxl)

test_data <- read_xlsx("C:/Users/elvir/anaconda3/data/IMPACT-Data-test-annexes/Annex 1 - REACH TEST DATA.xlsm")
village <-read_xlsx("C:/Users/elvir/anaconda3/data/IMPACT-Data-test-annexes/village.xlsm")
```









```{r}
test_data <- test_data %>%
  mutate(handwashing_check = case_when(handwashingfull=="No Handwashing facility" & `Frequency.respondant.report.handwhashing.a.day` =="7 times and more" ~ TRUE
                                        ,.default= FALSE ))
str(test_data)
```

# Question 13

You can also embed plots, for example:

```{r, echo=FALSE}
test_data <- test_data %>%
  mutate(diarrhea_under_5= case_when(InterviewID == "baseline1000" & diarrhea_under_5=="Yes"~"NA", .default= diarrhea_under_5)
         ,Improvedsanitationfacility = case_when(InterviewID=="uuid:fde636f2-60d5-4c61-bdb8-be783b8e207f" & Improvedsanitationfacility =="Improved toilet facility" ~"Unimproved toilet facility",.default = Improvedsanitationfacility ))

```

# Question 14

```{r}
test_data %>%
  group_by(data_collection_round, house_type) %>%
  summarise(n=n(), prop=n/sum(n)) %>%
  mutate(prop = n / sum(n))

```

# Question 15

```{r}
check_outliers <- function(column_to_check) {
  # Ignore missing values
  column_to_check <- column_to_check[!is.na(column_to_check)]
  
  # Calculate mean and standard deviation
  mean_value <- mean(column_to_check)
  sd_value <- sd(column_to_check)
  
  # Calculate the lower and upper bounds for outliers
  lower_bound <- mean_value - 3 * sd_value
  upper_bound <- mean_value + 3 * sd_value
  
  # Check for outliers
  outliers <- column_to_check < lower_bound | column_to_check > upper_bound
  
  return(outliers)
}


check_outliers(test_data$Number.adult.household.members.years.old)

```
# Question 16

```{r}

village$InterviewID <- village$uuid
test_data <- inner_join(test_data, village, by = c("InterviewID","data_collection_round"))

```


# Question 17

```{r}
test_data %>%
  group_by(data_collection_round, handwashingfull) %>%
  summarise(n=n()) %>%
  mutate(perc = n / sum(n) *100 )%>%
ggplot(aes(x = data_collection_round,y=perc, fill = handwashingfull)) +
geom_col() +
scale_fill_brewer() +
labs(x = "Round", y = "Percent",
title = "Percentage of type of hand washing facility for each round")
```
# Question 18

```{r}

# Function to count missing values in a vector
count_missing <- function(x) {
  sum(case_when(x=="NA" | is.na(x) == TRUE ~1,.default=0))
}

# Apply the function to each column of the data frame
missing_counts <- apply(test_data, 2, count_missing)

# Convert the result to a list
missing_list <- as.list(missing_counts)

missing_list
```
## Analysis

# Question 12


```{r}

test_data %>%
  mutate( safe_water = case_when( drinking_water_source %in% c("Protected dug well", "Piped water to yard or plot","Piped water into dwelling (house)","Bottled water", "Tube well or borehole","Public tap or standpipe", "Protected spring")~ "Improved water source"
                                  ,drinking_water_source %in% c("Cart with small tank or drum","Unprotected dug well", "Unprotected spring", "Surface water","Tanker-truck","Rainwater collection","Other")~"Unimproved water source")) %>%
  filter(is.na(safe_water)==0 & data_collection_round =="Baseline")%>%
  group_by( safe_water)%>%
  summarise(n=n())%>%
  mutate(perc = n/sum(n)*100)

```
The % of households having access to safe drinking  source of water is 92.0% at baseline

# Question 13

Access to safe water for single headed household at Baseline
```{r}

test_data %>%
  mutate( safe_water = case_when( drinking_water_source %in% c("Protected dug well", "Piped water to yard or plot","Piped water into dwelling (house)","Bottled water", "Tube well or borehole","Public tap or standpipe", "Protected spring")~ "Improved water source"
                                  ,drinking_water_source %in% c("Cart with small tank or drum","Unprotected dug well", "Unprotected spring", "Surface water","Tanker-truck","Rainwater collection","Other")~"Unimproved water source")) %>%
  filter(is.na(safe_water)==0 & data_collection_round =="Baseline" & single_headed_household =="Yes")%>%
  group_by( safe_water)%>%
  summarise(n=n())%>%
  mutate(perc = n/sum(n)*100)

```


Access to safe water for single headed household in End-line
```{r}

test_data %>%
  mutate( safe_water = case_when( drinking_water_source %in% c("Protected dug well", "Piped water to yard or plot","Piped water into dwelling (house)","Bottled water", "Tube well or borehole","Public tap or standpipe", "Protected spring")~ "Improved water source"
                                  ,drinking_water_source %in% c("Cart with small tank or drum","Unprotected dug well", "Unprotected spring", "Surface water","Tanker-truck","Rainwater collection","Other")~"Unimproved water source")) %>%
  filter(is.na(safe_water)==0 & data_collection_round =="End-line" & single_headed_household =="Yes")%>%
  group_by( safe_water)%>%
  summarise(n=n())%>%
  mutate(perc = n/sum(n)*100)

```
The access changed from 91% to 97%, so access to safe water increased between the period.


# Question 14



```{r}
test_data %>%
  group_by(data_collection_round, diarrhea_under_5) %>%
  summarise(n=n()) %>%
  mutate(perc = n / sum(n) *100 )%>%
  filter(diarrhea_under_5=="Yes")%>%
ggplot(aes(x = data_collection_round,y=perc)) +
geom_col() +
scale_fill_brewer() +
labs(x = "Round", y = "Percent",
title = "% of households with cases of diarrhea of children under 5 at the baseline and endline")
```

Yes, the graph is correct rounded off to the nearest digit

